---
title: "Exercise3"
author: "Giulio Vicentini"
date: "16/5/2020"
output: html_document
---

EX 1
determine the costant c
```{r}
f <- function(t) {
    val <- ifelse(t>1 & t<2, (t-1)*(2-t), 0)
    return(val)
}

c <- 1/integrate(f, 1, 2)$value
```

write the set of four R functions and plot the pdf and cdf
```{r}
dtolc <- function(t) {
    d <- ifelse(t>1 & t<2, c*(t-1)*(2-t), 0)
    return(d)
}

ptolc <- function(t) {
    p <- ifelse(t<=1, 0, ifelse(t>1 & t<2, -c*(t-1)^2*(2*t-5)/6, 1 )) 
    return(p)
}

rtolc <- function(n) {
    M <- 1.5
    u1 <- runif(n, 0, 1) 
    u2 <- runif(n, 0, 1)
    y <- ifelse(u2 * M < dtolc(u1+1), u1+1, NA)
    y_clean <- y[!is.na(y)]
    return(y_clean)
}

qtolc <- function(p) {
  q <- NULL
  j <- 1
  for (i in p) {
    f <- function(t) {ptolc(t)-i}
    root <- uniroot(f, lower=1, upper=2)$root
    q[j] <- root
    j <- j+1
  }
  return(q)
}

x <- seq(0, 3, 0.01)
plot(x, dtolc(x), type='l', main='PDF')
plot(x, ptolc(x), type='l', main='CDF')

sim <- rtolc(10000)
x <- seq(1, 2, 0.1)
hist(sim, col='gold')
curve(dtolc(x)*2000/3, add=TRUE)
```

evaluate the probability
```{r}
P_gt75 <- integrate(dtolc, 1.25, 2)$value

P_912 <- integrate(dtolc, 1.5, 2)$value
```
one point has negligible measure, hence P_90=P_120=0


EX 2
write the set of four R functions and plot the pdf and cdf
```{r}
dtires <- function(x) {
    d <- ifelse(x>= 1 & x<=2, 2/x^2, 0)
    return(d)
}

ptires <- function(x) {
  p <- ifelse(x<=1, 0, ifelse(x>1 & x<2, 2-2/x, 1))
  return(p)
}

rtires <- function(n) {
    u <- runif(n) 
    r <- -2/(u-2)
    return(r)
}

qtires <- function(p) {
  q <- -2/(p-2)
  return(q)
}

x <- seq(0, 3, 0.01)
plot(x, dtires(x), type='l', main='PDF')
plot(x, ptires(x), type='l', main='CDF')
```

evaluate the probability
```{r}
P_lt15 <- integrate(dtires, 1, 1.5)$value
```

sample 3000 random variables from the distribution and determine the mean value and the variance
```{r}
sim <- rtires(3000)
hist(sim, col='gold')
x <- seq(1, 2, 0.1)
curve(dtires(x)*500/2, add=TRUE)

E_X <- mean(sim)
E_X2 <- mean(sim^2)
VAR <- E_X2 - E_X^2

e_x <- integrate(function(x){x*dtires(x)}, 1, 2)$value
e_x2 <- integrate(function(x){x*x*dtires(x)}, 1, 2)$value
var <- e_x2 - e_x^2
```


EX 3
show, with a plot, that Markov’s inequality is is an upper bound to the following
distributions:
```{r}
#exponential distribution
k <- seq(0, 10, 0.01)
lambda <- 1
plot(k, 1-pexp(k, lambda), type='l', col='blue', main='Marokov’s inequality - exponential')
lim <- function(x) {1/(lambda*x)}
curve(lim, add=TRUE)

text(2.4, 0.52, 'E[X]/k')
text(1, 0.16, 'G(k)')

#uniform distribution
k <- seq(2, 6, 0.01)
a <- 3
b <- 5
plot(k, 1-punif(k, a, b), type='l', col='blue', main='Marokov’s inequality - uniform')
lim <- function(x) {(a+b)/(2*x)}
curve(lim, add=TRUE)

text(4.7, 0.9, 'E[X]/k')
text(4, 0.27, 'G(k)')

#binomial distribution
k <- seq(0, 10, 1)
n <- 1
p <- 1/2
plot(k, 1-pbinom(k, n, p), type='b', col='blue', main='Marokov’s inequality - binomial')
lim <- function(x) {n*p/x}
curve(lim, add=TRUE)

text(4, 0.15, 'E[X]/k')
text(1, 0.2, 'G(k)')

#poisson distribution
x <- seq(0, 10, 1)
lambda <- 1/2
plot(x, 1-ppois(x, lambda), type='b', col='blue', main='Marokov’s inequality - poisson')
k <- 2
lim <- function(x) {lambda/x}
curve(lim, add=TRUE)

text(2.4, 0.26, 'E[X]/k')
text(1, 0.03, 'G(k)')
```

EX 4
show, with a plot, that Chebyshev’s inequality is is an upper bound to the following
distributions:
```{r}
#normal distribution
k <- seq (0, 5, 0.01)
mu <- 3
sigma <- 5
int <- function(x) {pnorm(mu+x*sigma, mu, sigma) - pnorm(mu-x*sigma, mu, sigma)}
plot(k, int(k), type='l', col='blue', main='Chebyshev’s inequality - normal')
lim <- function(x) {1-1/x^2}
curve(lim, add=TRUE)
text(2, 0.4, '1-1/k^2')

#exponential distribution
k <- seq(0, 5, 0.01)
lambda <- 1
mu <- 1/lambda
sigma <- 1/lambda
int <- function(x) {pexp(mu+x*sigma, lambda) - pexp(mu-x*sigma, lambda)}
plot(k, int(k), type='l', col='blue', main='Chebyshev’s inequality - exponential')
lim <- function(x) {1-1/x^2}
curve(lim, add=TRUE)
text(2, 0.4, '1-1/k^2')

#uniform distribution
k <- seq(0, 5, 0.01)
a <- 1-sqrt(2)
b <- 1+sqrt(2)
mu <- (a+b)/2
sigma <- (b-a)/sqrt(12)
int <- function(x) {punif(mu+x*sigma, a, b) - punif(mu-x*sigma, a, b)}
plot(k, int(k), type='l', col='blue', main='Chebyshev’s inequality - uniform')
lim <- function(x) {1-1/x^2}
curve(lim, add=TRUE)
text(2, 0.4, '1-1/k^2')

#poisson distribution
k <- seq(0, 5, 1)
lambda <- 1/3
mu <- lambda
sigma <- sqrt(lambda)
int <- function(x) {ppois(mu+x*sigma, lambda) - ppois(mu-x*sigma, lambda)}
plot(k, int(k), type='b', col='blue', main='Chebyshev’s inequality - poisson')
lim <- function(x) {1-1/x^2}
curve(lim, add=TRUE)
text(2, 0.4, '1-1/k^2')
```

