---
title: "Exercise2_1"
author: "Giulio Vicentini"
date: "7/5/2020"
output: html_document
---

EX 1
```{r}
x <- c(15.58, 15.9, 16, 16.1, 16.2)
p1 <- c(0.15, 0.21, 0.35, 0.15, 0.14)
p2 <- c(0.14, 0.05, 0.64, 0.08, 0.09)

E1 <- sum(x*p1)
E2 <- sum(x*p2)

var1 <- sum(x^2*p1)-E1^2
var2 <- sum(x^2*p2)-E2^2
```


EX 2
plot histogram for 50 people
```{r}
lambda <- 1/30
n <- 50
oss <- rexp(n, lambda)
hist(oss, col='blue', xlab='x', ylab='dexp(y|lambda)', main='exponential distr. lambda=1/30')
```

probability that a person will wait for less than 10 minutes
```{r}
P_10 <- pexp(10, lambda)
```

average waiting time
```{r}
E_th <- 1/lambda
E_sper <- mean(oss)
```

probability for waiting more than one hour
```{r}
P_60 <- 1-pexp(60, lambda)
```


EX 3
assume lambda=1/3 and plot pdf and cdf
```{r}
lambda <- 1/3
x <- 0:10

dhist <- dpois (x, lambda)
phist <- ppois (x, lambda)

barplot(dhist, col='blue', ylim=c(0,1), names=0:10, xlab='x', ylab='dpois(x|lambda)', main='poisson distr. lambda=1/3')
barplot(phist, col='blue', ylim=c(0,1.2), names=0:10, xlab='x', ylab='ppois(x|lambda)', main='exponential cdf. lambda=1/3')
```

probabilty of at least one error
```{r}
P_1 <- 1 - ppois(0, lambda)
```


EX 4
follow the geometric distribution
```{r}
p <- 4/52
dgeom(10, p)
```


EX 5
load the data
```{r}
majors <- read.csv2("/home/giulio/Documenti/dataset/sindaciincarica.csv", skip=2)
```

plot sex amongst the majors
```{r}
library(dplyr)

sex <- count(majors, majors$sesso)
barplot(sex$n, col = c('mistyrose', 'lightblue'), ylim=c(0,7000), names=c('f', 'm'))
```

plot the number of towns grouped per province and per region
```{r}
prov <- count(majors, majors$codice_provincia)
barplot(prov$n, col = 'green', ylim=c(0,300), names=prov$`majors$codice_provincia`)

reg <- count(majors, majors$codice_regione)
barplot(reg$n, col = 'green', ylim=c(0,1600), names=reg$`majors$codice_regione`)
```

plot a distributions of the age of the mayors
```{r}
years <- count(majors, substr(majors$data_nascita, 7, 10))
barplot(years$n, col = 'green', ylim=c(0,300), names=years$`substr(majors$data_nascita, 7, 10)`)
```

plot a distribution of the time (in days) the mayor is in charge
```{r}
library(lubridate)

date <- count(majors, majors$data_elezione)
date_order <- date[order(as.Date(date$`majors$data_elezione`, '%d/%m/%Y')), ]
barplot(date_order$n, col = 'green', ylim=c(0,4000), names=today()-as.Date(date_order$`majors$data_elezione`, '%d/%m/%Y'), xlab='days', ylab='#majors')
```

how many majors are going to complete their mandate in 2020 and 2021
```{r}
count(majors, substr(majors$data_elezione, 7, 10)==2015)
count(majors, substr(majors$data_elezione, 7, 10)==2016)
```

